---
title: "White House Salary Gender Analysis"
date: '`r format(Sys.Date(), "%m-%d-%Y")`'
output: 
  github_document:
    html_preview: FALSE
  pdf_document:
    toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Comparing Salaries between male and female White House Staff Members

[American Progress Analysis](https://www.americanprogress.org/issues/economy/reports/2014/05/19/90039/explaining-the-gender-wage-gap/)

Recently I thought it would be interesting to see if men and women in
The White House receive equal pay and what those trends are over time.

For this analysis, I'll be relying heavily on the `gender` package in R
to assume the gender identity of each name based on the dataset provided.

I have not taken the time to spot check each time, but if you're
interested in teaming up and tackling it, 
[please let me know](http://www.twitter.com/chipoglesby)!

First let's simply start with visualizing pay between gender over time:

```{r genderLineChart, echo = FALSE, warnings = FALSE}
salaries %>% 
  filter(status == 'employee') %>% 
  group_by(year, gender) %>%
  summarize(medianSalary = median(salary)) %>% 
  ggplot(aes(year, medianSalary, color = gender)) +
  geom_line(size = 1.5) +
  scale_x_continuous(breaks = c(2001:as.numeric(format(Sys.Date(), "%Y")))) +
  ggtitle("Median Pay Over Time")
```

As we've seen in earlier in my first analysis, median salaries are on
the rise year over year.

```{r genderBoxplot, echo = FALSE, warnings = FALSE}
salaries %>% 
  filter(status == 'employee') %>% 
  group_by(gender) %>% 
  ggplot(aes(gender, salary, fill = gender)) +
  geom_boxplot() +
  ggtitle('Boxplot: Salaries of Men vs Women') +
  facet_wrap(~year)
```