---
title: "Staff Term Length"
date: '`r format(Sys.Date(), "%m-%d-%Y")`'
output: 
  github_document:
    html_preview: FALSE
  pdf_document:
    toc: false
---

```{r setup, include=FALSE}
library(tidyverse)
library(magrittr)
library(knitr)
options(digits = 15)
library(scales)
```

## How long does the average employee stay on staff?

Let's start by looking at the average employee length of stay between
`r min(salaries$year)` and `r max(salaries$year)`.
 
```{r, averageStay, echo = FALSE}
yearCount %>% 
  ggplot() +
  geom_histogram(aes(years), bins = 50)
```

It looks like the majority of employees will only work one year and there
also appears to be a handful who have worked as many as 
`r max(longestStaff$years)` years. We'll investigate that later.

Next, let's take at look at the top positions with the highest turnover.
These positions have the highest count of employees who only
work one year.

```{r, staffTurnOver, echo = FALSE}
yearCount %>% 
  filter(years == 1) %>% 
  inner_join(salaries, 'employee_name') %>% 
  group_by(position) %>% 
  summarize(count = n()) %>%
  arrange(desc(count)) %>% 
  top_n(10, count) %>% kable()

yearCount %>% 
  filter(years == 1) %>% 
  inner_join(salaries, 'employee_name') %>% 
  group_by(position) %>% 
  summarize(count = n()) %>%
  arrange(desc(count)) %>% 
  top_n(10, count) %>% 
  ggplot(aes(position, count)) +
  geom_bar(stat = 'identity') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Longest Employed Staff

One interesting nugget of data that I've found during my analysis
is that some staffers have worked as many as 
`r max(longestStaff$years)` years in the White House. I thought
it would be interesting to look more at these people and staff 
overall.

Here I am going to make the assumption that the people listed
below are the same people, because they've held the same or similiar
positions over time.

```{r, staffTable, echo = FALSE}
longestStaff %>% 
  kable()
```

That's interesting, right? `r nrow(longestStaff)` people have been working in
The White House for `r max(longestStaff$years)` years now. I wonder
what they do there?

Let's take a deeper dive into the data:

```{r, deepDive, echo = FALSE}
longestStaff %>% 
  inner_join(salaries, "employee_name") %>% 
  select(employee_name, year, position, president) %>% 
  arrange(employee_name, year) %>% 
  kable()
```

## Visualizing Employee Tenure

It would be interesting to know if those `r nrow(longestStaff)` are
compensatated properly for their tenure while working at The White House?

```{r, staffPlots, echo = FALSE}
# Plots for Longest Employeed Staff
for (i in 1:length(longestStaff$employee_name)) {
  print(salaries %>% 
          filter(grepl(longestStaff$employee_name[[i]], employee_name)) %>% 
          ggplot(aes(year, salary)) +
          geom_line(size = 1.5) +
          ggtitle(tools::toTitleCase(longestStaff$employee_name[[i]])) +
          scale_y_continuous(labels = comma) +
          scale_x_continuous(breaks = c(2001:as.numeric(format(Sys.Date(), "%Y")))) +
          theme(axis.text.x = element_text(angle = 45, hjust = 1)))
}
```